Printing in a Periphery: a Quantitative Study of Finnish Knowledge Production, 1640-1828

### 2016-07-06 inst/examples/20160715-slides-Krakow.Rmd
a
s

@@

Idea: to study early modern knowledge production in general.

@@

What is important: to clean up the data (90% of the work) and understand how particular data collections have been formed.

Fail: take worldcat or similar combination of datasets and analyse it without cleaning it first.

@@

## Data and method

@@

- Kungliga (Royal Swedish Library's metadata), n-number of documents

@@

@@

## Argument and findings

@@

- Academy of Turku as part of the Swedish University system is the centre in Finland.

Developments that start after "the rock bottom" of 1700-1721

@@

3) Economy (and related subject fields) published in Finnish

4) Enlightenment in Sweden reflected in octavo sized book publishing

@@

## Peaks and drops – role of publishers and contingency

@@

## Comparing publishing activity

@@

## Unique publishers in top publication places over time

@@

@@

## Title count per top publisher

@@

## Title count per top publisher

```{r publishers3-finland-kungliga, echo=FALSE, message=FALSE, warning=FALSE, cache=FALSE, fig.height=7, fig.width=13, out.width="150px", fig.show="hold"}
# https://github.com/rOpenGov/fennica/blob/master/inst/examples/figure_201606_Krakow/publishers3-finland-1.png
catalogue = "Kungliga"
place <- "Turku"
df = df0

# TITLE COUNT
theme_set(theme_bw(20))
p <- ggplot(npub, aes(x = publication_decade, y = n)) +
       geom_bar(stat = "identity", position = "stack", aes(fill = publisher)) + 
       xlab("Publication year") +
       ylab("Title count (n)") +        
       ggtitle(paste("Title count per publisher (", place, "/", catalogue, ")", sep = ""))
print(p)
```

## Publisher paper vs. title count

Paper per title count ratio per publisher. [nähdään voidaanko sanoa onko joku selkeästi asetusten julkaisija ja toisella myös kirjoja (tai asetuskokoelmia).]

@@

## Drop 1700

Publisher title count changes around 1700 drop in Turku, the most significant publishing place at the time:

@@

## Gävle in Riksdagar

@@

Gävle (gray) during riksdagar (Kungliga)

@@

# Language

## Book printing in Finland / Fennica

@@

## Book printing in Finland / Kungliga

@@

## Same with percentages

@@

@@

## Religious literature

Devotional literature (catechisms, hymns, prayer collections, etc.) is known to be the most important form of literature denoting the advancement of reading in Finland in nineteenth century. But when does it actually showup as a relevant category?

@@

### BOOKINFINLAND DONE3

Temporary sermons played a crucial role in early modern knowledge production, but it has not been clearly analysed in what language they were in Finland (and Sweden).

@@

### BOOK IN FINLAND 4

Combination of devotional literature and temporary sermons gives us the following graphs.

@@

# University teaching

### topics 1

Printing of theology related topics in different languages in Turku.

@@

### topics 2

Printing of history and related topics in different languages in Turku.

@@

### topics 3

Printing of philosophy in different languages in Turku.

@@

### topics 4

Printing of natural science in different languages in Turku.

@@

### topics 5

Printing of economy and related topics in different languages in Turku.

@@

# Enlightenment and reading

Rise of the octavo sized book as vehicle of Enlightenment (and proof of reading).

Paper consumption for different document formats over time. Each point represents a decade. Loess smoothing.

@@

## Turku

Effects of Enlightenment: decline of metaphysics in favour of social and economic theory and the great victory of reading is visible in the rise and rise of octavo book.

@@

## Thanks !

>>>>

[Intro 1/3]

@@

Idea: to study early modern knowledge production, all of it (1470-1830)

@@

What is important: to clean up the data (90% of the work) and to take into account how particular data collections (Finnish National Bibliography, English Short-Title Catalogue etc.) have been formed.

How to fail: take worldcat or similar combination of datasets and analyse it as a whole without cleaning it first.

@@

[Intro 2/3]

### Data and method

@@

- Kungliga (National Library of Sweden's library catalogue), n-number of documents

@@

- Benefit no. 1: once the work has been done, the data is available for everyone to use and to enrich it. The ecosystem is constructed with even unforeseen changes in mind.

@@

[Intro 3/3]

### Argument and findings

@@

- Academy of Turku as part of the Swedish University system is the centre of Finnish knowledge production.

Enlightenment in Finland after "the rock bottom" of 1700-1721

@@

3) Economy and social science published in Swedish language

--> Deep roots of the diversity of "Finnishness" clearly reflcted in knowledge production, 1640-1828.

@@

[Big Picture 1/4]

Kartta jossa näkyy Turku osana Ruotsin valtakuntaa tähän (Janilla ideoita manuskriptin puolella, mun mielestä ne on hyviä tästä). Voisi miettiä, että tämä kartta-slide korvaisi sitten tuon alla tulevan "Comparing publishing activity in Swedish University town".

[Big Picture 2/4]

### Peaks and drops – role of publishers and contingency

@@

[Big Picture 3/4]

### Comparing publishing activity in Swedish University towns

@@

[Big Picture 4/4]
### Rise of the octavo sized book as vehicle of Enlightenment

Paper consumption for different document formats over time.

### Publishers

[Publishers 1/6]

Turku in Sweden -slide (Janin ideoima tulisi tähän)

[Publishers 2/6]

### Development of number of publishers in Finland

@@

[Publishers 3/6]

### Development of number of publishers in Swedish University towns

@@

[Publishers 4/6]

### Numbers of publications of main publishers in Turku

@@

[Publishers 5/6]

### Comparing how much paper different publishers used on average in their publications in Turku

@@

[Publishers 6/6]

### Rock Bottom of the Great Northern War (1700-1721)

Publisher title count changes around 1700 with a nearly detrimental drop in Turku, the most significant publishing place at the time.

@@

[Swedish age of Freedom and abolishment of censorship 1/2]

### Gävle in Riksdagar

@@

Gävle (gray) during riksdagar (Kungliga) TÄÄ EI TOIMI NYT OLLENKAAN KUN ON AIKA PIENI TOI NOUSU NÄKYVISSÄ TÄSSÄ. MENEEKÖ OIKEIN?

@@

[Swedish age of Freedom and abolishment of censorship 2/2]

Tämän kohdalta olisi tärkeä, että mukana myös joku slide siitä 1760-1770 ajasta. Pitää myös miettiä mihin kohtaan esitystä tämä osio istuu.

### Languages

[Languages 1/3]

### Development of use of different languages in Finnish National Bibliography

@@

[Languages 2/3]

### Development of use of different languages in Kungliga Biblioteket's Swedish data

@@

[Languages 3/3]

### Percentages of use of different languages in Finnish and Swedish data across time

TÄMÄ PITÄÄ PÄIVITTÄÄ NIIN ETTÄ ON SE RANSKA JA VENÄJÄ MUKANA!

@@

}
```
---

### Religious texts in Finland

@@

[Religious texts in Finland 1/3]

### Devotional literature (catechisms, hymns, prayer collections, etc.)

- known as an form of literature denoting the advancement of reading in Finland in the nineteenth century. Yet, it has been unclear when does it show up as a relevant category.

@@

[Religious texts in Finland 2/3]

### Temporary sermons in Latin and Swedish

- Temporary sermons played a crucial role in early modern knowledge production, but it has not been clearly analysed in what language they were in Finland (and Sweden).

@@

[Religious texts in Finland 3/3]

### Combining devotional literature and temporary sermons

@@

### University curriculum

[University curriculum 1/6]

### Printing of theology in different languages in Turku

@@

[University curriculum 2/6]

### Printing of history in different languages in Turku

@@

[University curriculum 3/6]

### Printing of philosophy in different languages in Turku

@@

[University curriculum 4/6]

### Printing of natural science in different languages in Turku

@@

[University curriculum 5/6]

### Printing of economy in different languages in Turku

@@

[University curriculum 6/6]

### Effects of the Enlightenment on the curriculum in Turku

- decline of metaphysics in favour of social and economic theory is clear in Turku
- the Enlightenment literature that people read in Finland was mainly printed in Stockholm and on the continent.

Tähän top-auktorilistat Fennica ja Kungliga.

@@

[Conclusion 1/2]

### Finland in the Swedish context

- Dramatic events regarding the diversity of "Finnishness" take place in the early nineteenth century. Finland becomes part of Russia and Turku is stripped off of its former glory... But the analysis of Finnish knowledge production 1828-2016 is for another day.

@@

[Conclusion 2/2]

### Thanks !
